---
output: github_document
---
<!-- .md is generated from .Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, message = FALSE, include=FALSE}
library(dplyr)
library(sastats)
```

# Overview

Package sastats includes a few convenience functions for estimating standard errors. This is typically in the context of survey sampling error. These are very simple computations, and the functions are just thin wrappers for a bit of code.

Using sample survey data:

```{r}
library(dplyr)
library(sastats)
data(svy)

activity <- left_join(svy$act, select(svy$person, Vrid, weight), by = "Vrid")
glimpse(activity)
```

## Proportion

Looking at participation rate:

```{r}

```

#### Computation

```{r}
error_se_prop
```

## Mean

Looking at days or participation:

```{r}
activity %>%
    group_by(act) %>% 
    summarise(
        avgdays = mean(days, na.rm = TRUE),
        se = error_se_mean(days, na.rm = TRUE)
)
```

#### Computation

```{r}
error_se_mean
```

