---
output: github_document
---
<!-- .md is generated from .Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, message = FALSE, include=FALSE}
library(dplyr)
library(sastats)
```

# Overview

Package sastats includes a few convenience functions for estimating survey sampling errors. These are simple computations, and the functions are just thin wrappers for a bit of code. Relevant measures:

- standard error of the mean
- standard error of a proportion
- margin of error

Note that [package survey](https://cran.r-project.org/web/packages/survey/index.html) provides a much more comprehensive approach to survey-based calculations (errors, weighting, etc.), but I haven't used it.

#### Sample Data

```{r}
library(dplyr)
library(sastats)
data(svy)

activity <- left_join(svy$act, select(svy$person, Vrid, weight), by = "Vrid")
glimpse(activity)
```

## SE Mean

Looking at days of participation:

```{r}
days <- activity %>%
    group_by(act) %>% 
    summarise(
        avgdays = weighted.mean(days, weight, na.rm = TRUE),
        se = error_se_mean(days, na.rm = TRUE)
)
days
```

## SE Proportion

Looking at participation rate:

```{r}
rate <- activity %>%
    group_by(act, part) %>%
    summarise(n = n(), wtn = sum(weight)) %>%
    mutate(
        n = sum(n), 
        rate = wtn / sum(wtn),
        se = error_se_prop(rate, n)
    ) %>%
    filter(part == "Checked")
rate
```

## Margin of Error

These are useful for reporting confidence intervals. Note that the weighting that was done to this dataset produces a "design effect" that inflates the margin of error. I know the design effect is 1.15 for this dataset (from the `rake_weight()` procedure):

```{r}
deff <- 1.15
mutate(rate, me = error_me(se) * deff)

mutate(days, me = error_me(se) * deff)
```
