---
output: github_document
---
<!-- .md is generated from .Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, message = FALSE, include=FALSE}
library(ggplot2)
library(dplyr)
library(sastats)
```

# Overview

Package sastats includes a few convenience functions for estimating survey sampling errors. These are simple calculations, and the functions are just thin wrappers for a bit of code. Relevant measures:

- Standard error of the mean
- Standard error of a proportion
- Margin of error (for confidence intervals)

### Notes

[Package survey](https://cran.r-project.org/web/packages/survey/index.html) provides a much more comprehensive approach to survey-based calculations (errors, weighting, etc.). I've tended toward the light-weight approach outlined here, but it could be worth looking into if we are doing alot of survey analysis in R.

Caculating errors for multi-estimate metrics requires [error propagation](https://en.wikipedia.org/wiki/Propagation_of_uncertainty) (e.g., total days which depends on participation rate & average days) . I haven't implemented these computations here, but existing packages address this need. Package  [propagate](https://cran.r-project.org/web/packages/propagate/index.html) is one I've used, but have found tricky to implement. Package  [errors](https://github.com/r-quantities/errors) is newer and appears more straightforward (although I haven't tested it).

#### Example Data

For demonstration, package sastats includes a survey dataset with annual participation metrics for 9 outdoor recreation activities:

```{r}
library(dplyr)
library(sastats)
data(svy)

activity <- left_join(svy$act, select(svy$person, Vrid, weight), by = "Vrid")
glimpse(activity)
```

## SE Mean

Looking at days of participation:

```{r}
days <- activity %>%
    group_by(act) %>% 
    summarise(
        avgdays = weighted.mean(days, weight, na.rm = TRUE),
        se = error_se_mean(days, na.rm = TRUE)
)
days
```

## SE Proportion

Looking at participation rate:

```{r}
rate <- activity %>%
    group_by(act, part) %>%
    summarise(n = n(), wtn = sum(weight)) %>%
    mutate(
        n = sum(n), 
        rate = wtn / sum(wtn),
        se = error_se_prop(rate, n)
    ) %>%
    filter(part == "Checked")
rate
```

## Margin of Error

These are useful for reporting confidence intervals. Note that the dataset weighting produces a "design effect" which inflates the margin of error. I know the design effect is 1.15 for this dataset, based on the summary output produced by the `sastats::rake_weight()` procedure.

```{r}
deff <- 1.15
rate <- mutate(rate, me = error_me(se) * deff, lower = rate - me, upper = rate + me)
days <- mutate(days, me = error_me(se) * deff, lower = avgdays - me, upper = avgdays + me)

library(ggplot2)
ggplot(rate, aes(act, rate)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper)) +
  ggtitle("Participation Rate Estimates (with 95% CIs)")

ggplot(days, aes(act, avgdays)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper)) +
  ggtitle("Average Days (per participant) Estimates (with 95% CIs)")
```
