---
output: github_document
---
<!-- .md is generated from .Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, message = FALSE, include=FALSE}
library(dplyr)
library(sastats)
```

# Overview

Rake weighting in R is fairly straightforward using [package anesrake](https://cran.r-project.org/web/packages/anesrake/index.html). The key aspect is getting the two input datasets into the right format:

- The **target population dataset** needs to be a named list (one element per demographic variable) of target distributions. The `weights::wpct()` function helps with this.

- The **survey dataset** needs to be a data frame with a unique ID variable (e.g., Vrid), and one variable for each demographic measure.

Importantly, demographic variables must be stored either as factors (or numerics) and these much match between the two datasets. This is probably confusing, but the examples below should help illustrate. There is also an AS/HS implementation on Github: [rakewt-ashs](https://github.com/southwick-associates/rakewt-ashs).

#### Example Data

For demonstration, package sastats includes survey and population datasets which include several demographic variables:

```{r}
library(dplyr)
library(sastats)

data(svy, pop)
svy <- select(svy$person, -weight)

# survey to be weighted
glimpse(svy)

# target population (outdoor recreationists) from a genpop survey
glimpse(pop)
```

Importantly, the demographic variables have been encoded in the same way for the 2 datasets (as factors in this case):

```{r}
wtvars <- setdiff(names(svy), "Vrid")

sapply(wtvars, function(x) all(levels(svy[[x]]) == levels(pop[[x]])))
```

## Population Distributions

The easiest way to see the format needed for the target population is to run `weights::wpct()` on a demographic variable. In this example, we are defining the target population (outdoor recreationists) using another general population survey dataset. The genpop distribution for age can be readily pulled from the pop dataset, although we need to ensure we use the `stwt` variable from that survey:

```{r}
weights::wpct(pop$age_weight, pop$stwt)
```

The `weights::wpct()` function returns a vector, but we need a list of vectors (one element per demographic variables of interest). A straightforward method involves looping over the demographic variables with `sapply()`:

```{r}
pop_target <- sapply(wtvars, function(x) weights::wpct(pop[[x]], pop[["stwt"]]))
pop_target
```

This also provides a convenient method for comparison with the survey dataset to be weighted:

```{r}
sapply(wtvars, function(x) weights::wpct(svy[[x]]))
```

## Rake Weighting

We can now run the rake weighting procedure. A summary will be printed  with some useful statistics (including design effect).

```{r}
svy <- rake_weight(svy, pop_target, "Vrid")
```

